---
title: "Lab 9 - Data Transformation"
author: "Victoria Sass"
date: "October 29, 2017"
output:
  pdf_document: default
  html_document: default
---

Using your own dataset (which may include more than one table) carry out the following data cleaning steps. Knit together the PDF document and commit both the Lab 9 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository. 

git add .
git commit -m ""

1. In addition to simply naming variable names in select you can also use `:` to select a range of variables and `-` to exclude some variables, similar to indexing a `data.frame` with square brackets. You can use both variable's names as well as integer indexes.


  a. Use `select()` to print out a tbl that contains only the  first 3 columns of your dataset, called by name.
  b. Print out a `tbl` with the last 3 columns of your dataset, called by name.
  c. Find the most concise way to select the first 3 columns and the last 3 columns by name. 

```{r}
library(tidyverse)
setwd("/Users/marakage/Desktop/Honors/STAT/Labs/Lab-9") 
load("lab8.RData")

#a.
select(barpop, POPGROUP.id, Asiansubgroup, subpoptotal)

#b.
select(barpop, Asiansubgroup, subpoptotal, subgroup_prop)

#c. besides calling the specific subroup name as above, maybe? Select(barpop, num_range(barpop, 1:3)) or head() and tail()

```



2. `dplyr` comes with a set of helper functions that can help you select groups of variables inside a `select()` call:
- `starts_with("X")`: every name that starts with "X",
- `ends_with("X")`: every name that ends with "X",
- `contains("X")`: every name that contains "X",
- `matches("X")`: every name that matches "X", where "X" can be a regular expression,
- `num_range("x", 1:5)`: the variables named x01, x02, x03, x04 and x05,
- `one_of(x)`: every name that appears in x, which should be a character vector.

```{r}
library(dplyr)

select(AsianAlone, starts_with("A")) 
select(AsianAlone, ends_with("n"))
select(AsianAlone, contains("a"))
select(AsianAlone, matches("c"))
select(AsianAlone, num_range("019", 1:5))
Asiansubgroups <- c("GEO.id","GEO.id2")
select(AsianAlone, one_of(Asiansubgroups))
       
```


Pay attention here: When you refer to columns directly inside `select()`, you don't use quotes. If you use the helper functions, you do use quotes.

  a. Use `select()` and a helper function to print out a `tbl` that selects only variables that contain a specific character string.
  
  b. Use `select()` and a helper function to print out a `tbl` that selects only variables that start with a certain letter or string of letters.
```{r}
#a.
select(AsianAlone, contains("laotian"))
#b.
select(AsianAlone, starts_with("l"))

```



4. Are there any mutations you wish to carry out on your data (i.e. new variables you wish to create based upon the values of already existing variables)? If so, describe what they are and what you will name them.

For the barchart, I will need to create proportion column with the proportion of each subgroup population in comparison with total population in King County. It will be named "prop". The mutation will happen as I first need to summarize the total count for each subgroup, and remove the Asian Alone in combination with as it is not clear how the individuals of this category were grouped, mutate to create a new total population column, adding the proportional relation between subgroup and total population. 


5. You can use mutate() to add multiple variables at once. To create more than one variable, place a comma between each variable that you define inside `mutate()`.
  
  
  a. Carry out any and all of the mutations you wish to perform on your dataset and print the results to the console.
  
  
```{r}
Data_Asian_Alone %>%
group_by(POPGROUP.id) %>% # Group by ethnicity/subgroup
  summarise(Total_pop = sum(Count)) %>% #Sum within subgroup
  filter(POPGROUP.id != 031) %>% #Remove asian and other
  mutate(Total_Asian = sum(Total_pop), #Create total asian pop variable
         proportion = Total_pop/Total_Asian) # Creating proportion of subgroup of total Asians
```

6. `R` comes with a set of logical operators that you can use inside `filter()`:
- `x < y`, `TRUE` if `x` is less than `y`
- `x <= y`, `TRU`E if `x` is less than or equal to `y`
- `x == y`, `TRUE` if `x` equals `y`
- `x != y`, `TRUE` if `x` does not equal `y`
- `x >= y`, `TRUE` if `x` is greater than or equal to `y`
- `x > y`, `TRUE` if `x` is greater than `y`
- `x %in% c(a, b, c)`, `TRUE` if `x` is in the vector `c(a, b, c)`

  a. What are some potential subsets of your data that seem interesting and worth investigation to you? 
  
  I will be investigating the proportionality of the population of each Asian subgroup in relationship with total population of King County for the first visual, and faceting 11 maps depicting each Asian subgroup, plus Asian alone group and total population to geographicly represent each group by tract representation for the second visual representation.

  
  b. Use at least two of the logical operators presented above to print these subsets of your data.
    ```{r}

    testAsian <- Data_Asian_Alone %>%
      filter (Count  >= 100 & Count <= 200)
    
  
    ```
    
     
7. `R` also comes with a set of boolean operators that you can use to combine multiple logical tests into a single test. These include `&` (and), `|` (or), and `!` (not). Instead of using the `&` operator, you can also pass several logical tests to `filter()`, separated by commas. `is.na()` will also come in handy.   
```{r}

  #filter(Data_Asian_Alone, Count > 100 & POPGROUP.id == "015")
Data_Asian_Alone %>%
  filter(!is.na(Count))
         
```

  a. Use `R`'s logical and boolean operators to select just the rows in your data that meet a specific boolean condition. 
  b. Print out all of the observations in your data in which none of variables are `NA`.


```{r}

is.na(AsianAlone) 
anyNA(AsianAlone, recursive= TRUE)

#there are no observations NA in AsianALone table



```



8. `arrange()` can be used to rearrange rows according to any type of data. If you pass `arrange()` a character variable, for example, `R` will rearrange the rows in alphabetical order according to values of the variable. If you pass a factor variable, `R` will rearrange the rows according to the order of the levels in your factor (running `levels()` on the variable reveals this order).

```{r}
testAsian_b <- 
  select(AsianAlone, starts_with("A")) %>%
         arrange(Asiansubgroups)

```


By default, `arrange()` arranges the rows from smallest to largest. Rows with the smallest value of the variable will appear at the top of the data set. You can reverse this behavior with the `desc()` function. `arrange()` will reorder the rows from largest to smallest values of a variable if you wrap the variable name in `desc()` before passing it to `arrange()`.

  a. Which variable(s) in your dataset would be logical to arrange your data on? Explain your reasoning.
  
  From the main AsianAlone dataset,it is not important to arrange the data in a specific order as it is a categorical descriptive analysis of each subgroup and its population represented. 
  
  b. Arrange your data by this/these variables and print the results. 
  
  NA



9. You can use any function you like in `summarise()` so long as the function can take a vector of data and return a single number. `R` contains many aggregating functions, as `dplyr` calls them:
- `min(x)` - minimum value of vector `x`.
- `max(x)` - maximum value of vector `x`.
- `mean(x)` - mean value of vector `x`.
- `median(x)` - median value of vector `x`.
- `quantile(x, p)` - pth quantile of vector `x`.
- `sd(x)` - standard deviation of vector `x`.
- `var(x)` - variance of vector `x`.
- `IQR(x)` - Inter Quartile Range (IQR) of vector `x`.
- `diff(range(x))` - total range of vector `x`.

  a. Pick at least one variable of interest to your project analysis.
  b. Print out at least three summary statistics using `summarise()`. 
  
  
```{r}
as.numeric(AsianAlone$D001)
testasian_c <- summarise(AsianAlone, 
                         mean(as.numeric(D001), na.rm = TRUE))
                        
```



10. `dplyr` provides several helpful aggregate functions of its own, in addition to the ones that are already defined in `R`. These include:
- `first(x)` - The first element of vector `x`.
- `last(x)` - The last element of vector `x`.
- `nth(x, n)` - The nth element of vector `x`.
- `n()` - The number of rows in the data.frame or group of observations that `summarise()` describes.
- `n_distinct(x)` - The number of unique values in vector `x`.

Next to these `dplyr`-specific functions, you can also turn a logical test into an aggregating function with `sum()` or `mean()`. A logical test returns a vector of TRUE's and FALSE's. When you apply `sum()` or `mean()` to such a vector, `R` coerces each `TRUE` to a `1` and each `FALSE` to a `0`. `sum()` then represents the total number of observations that passed the test; `mean()` represents the proportion.

  a. Print out a summary of your data using at least two of these `dplyr`-specific aggregate functions. 
  b. Why did you choose the ones you did? What did you learn about your data from these summaries?
  


11. You can also combine `group_by()` with `mutate()`. When you mutate grouped data, `mutate()` will calculate the new variables independently for each group. This is particularly useful when `mutate()` uses the `rank()` function, that calculates within-group rankings. `rank()` takes a group of values and calculates the rank of each value within the group, e.g.

`rank(c(21, 22, 24, 23))`

has the output 

`[1] 1 2 4 3`

As with `arrange()`, `rank()` ranks values from the smallest to the largest.

  a. Using the `%>%` operator, first group your dataset by a meaningful variable, then perform a mutation that you're interested in. 
  
  
  b. What do the results tell you about different groups in you data?
  
```{r}
#a.
  barpop <- mergedata %>%
  mutate(Asiansubgroup = `POPGROUP.display-label`) %>%
  group_by(POPGROUP.id, Asiansubgroup) %>%
  summarise(subpoptotal = sum(Count)) %>%
  mutate(subgroup_prop = round(subpoptotal/sum(data_subset_total$TotalCount), 5)) %>%
  filter(POPGROUP.id != "031")

#b. The results show the proportion of each Asian subgroup with the total population in King County. The numbers/proportions are very low in general, but particularly low for a few subgroups. But overall there is a significant proportional difference among the Asian subgroups.  


```

12. The exercises so far have tried to get you to think about how to apply the five verbs of `dplyr` to your data. 
  
  a. Are there any specific transformations you want to make to your data? What are they and what aspect of your research question will they help you to answer?
  
  The transformations have been divided in two groups. One is the above mentioend population representation of each Asian subgroup in relationship with the total population.  The other is the representation of eah subgroup in each tract in King County. With both visual representations, I hope to convey the diversity  of the Asian population, emphasizing the large gap in population representation, and other socioeconomic variables, to debunk pervasive narratives that group together "Asians", and outcomes in narrative, discourse and policies that affect particularly the subgroups mariginalized voices. 
  

  b. In a code chunk below, carry out all the data transformations you wish to perform on your data. Utilize the `%>%` operator to tie multiple commands together and make your code more readable and efficient. Remember to comment your code so it is clear why you doing things a certain way. 
  
  
  
```{r}
#this is the same code performed in Lab8

  AsianAlone <- AsianAlone[-1,]
prop_data <- AsianAlone %>%
  mutate(Count = as.numeric(as.character(D001)), 
         POPGROUP.id = as.numeric(as.character(POPGROUP.id))) 
# Make new count variable which is numeric version of D001
  
prop_data %>%
  group_by(POPGROUP.id) %>% # Group by ethnicity/subgroup
  summarise(Total_pop = sum(Count)) %>% #Sum within subgroup
  filter(POPGROUP.id != 031) %>% #Remove asian and other
  mutate(Total_Asian = sum(Total_pop), #Create total asian pop variable
         proportion = Total_pop/Total_Asian) # Creating proportion of subgroup of total Asians
  
hist(data_subset_cambodian$Count)
#Changing factors to numeric on D001 column/population

pop.chinese <- sum(as.numeric(data_subset_chinese$D001), na.rm=FALSE)
pop.cambodian <- sum(as.numeric(data_subset_cambodian$D001), na.rm=FALSE) 
pop.japanese <- sum(as.numeric(data_subset_japanese$D001), na.rm=FALSE) 
pop.korean <- sum(as.numeric(data_subset_korean$D001), na.rm=FALSE)
pop.laotian <- sum(as.numeric(data_subset_laotian$D001), na.rm=FALSE)
pop.hmong <- sum(as.numeric(data_subset_hmong$D001), na.rm=FALSE)
pop.taiwanese <- sum(as.numeric(data_subset_taiwanese$D001), na.rm=FALSE) 
pop.vietnames <- sum(as.numeric(data_subset_vietnamese$D001), na.rm=FALSE)
pop.filipino <- sum(as.numeric(data_subset_filipino$D001), na.rm=FALSE)

pop <- c(pop.chinese, pop.cambodian, pop.japanese, pop.korean, pop.laotian,
                 pop.hmong, pop.taiwanese, pop.vietnames, pop.filipino) %>% as.data.frame() %>% t()

colnames(pop) <- c("Chinese", "Cambodian", "Japanese", "Korean", "Laotian",
                   "Hmong", "Taiwanese", "Vietnames", "Filipino")


asian.pop <- data.frame("Chinese", pop.chinese, stringsAsFactors = FALSE)
names(asian.pop) <- c("Asian Population", "Total Population")
asian.pop[nrow(asian.pop) + 1,] = c("Cambodian", pop.cambodian)

```

  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
