---
title: "Lab 8"
author: "Mara Kage"
date: "October 27, 2017"
output:
  pdf_document: default
  html_document: default
---

Using your own dataset (which may include more than one table) carry out the following data cleaning steps. Knit together the PDF document and commit both the Lab 8 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository. 

**Before you begin: as many of you have large datasets, you're going to want to select only the variables you're interested in utilizing for this project (ideally no more than twenty columns but perhaps much smaller) so you don't have R Studio's memory working on the entire dataset. The example code provided below can be modified to allow you to subset your data to only the variables you wish to use. First, read in your complete dataset and save it as** `data`. **Then, add the names of the variables you wish to use for your poster project to the** `select` **function, separated by commas. Run the two lines of code to save this new, smaller version of your data to** `data_subset`. **Use this smaller dataset to complete the rest of the lab**

```{r}
setwd("/Users/marakage/Desktop/Honors/STAT/Labs")
#install.packages("dplyr")
library(dplyr)
library(tidyverse)

# Read in your data with the appropriate function
#Download main tables from Census
AsianAlone <- read.csv("DEC_10_SF2_PCT1_with_ann.csv")
Education <- read.csv("../Datasets/Education/ACS_15_5YR_S1501_with_ann.csv")
Totalpopulation <- read_csv("../Datasets/TotalPopulation/ACS_10_5YR_B01003_with_ann.csv",skip=0)

#created data subsets of each Asian subgroup by tract
data_subset_cambodian <- AsianAlone %>%
  filter(POPGROUP.id == "015") %>%
  mutate(Count = as.numeric(as.character(D001))) %>% # Conversion for factor
  dplyr::select(POPGROUP.display-label, GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_chinese <- AsianAlone %>%
  filter(POPGROUP.id == "016") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_japanese <- AsianAlone %>% 
  filter(POPGROUP.id == "022") %>% 
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_korean <- AsianAlone %>%
  filter(POPGROUP.id == "023") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_laotian <- AsianAlone %>%
  filter(POPGROUP.id == "024") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_hmong <- AsianAlone %>%
  filter(POPGROUP.id == "020") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_filipino <- AsianAlone %>%
  filter(POPGROUP.id == "019") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_taiwanese <- AsianAlone %>%
  filter(POPGROUP.id == "018") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_vietnamese <- AsianAlone %>%
  filter(POPGROUP.id == "029") %>%
  mutate(Count = as.numeric(as.character(D001))) %>%
  select(`POPGROUP.display-label`,GEO.id, GEO.id2, `GEO.display-label`, Count)

data_subset_total <- Totalpopulation %>% 
  filter(GEO.id != "Id") %>%
  mutate(TotalCount = as.numeric(as.character(HD01_VD01))) 

#transform from factor to numeric
Data_Asian_Alone <- AsianAlone %>%
  filter(GEO.id != "Id") %>%
  filter(POPGROUP.id != "012" & POPGROUP.id != "031") %>% #removing the Asian Alone and Asian in combination groups
  mutate(Count = as.numeric(as.character(D001))) #creating a new column for the population number as numeric vs character

#barchart
mergedata <- data_subset_total %>% 
  left_join(Data_Asian_Alone, by="GEO.id2") %>%
  select(-GEO.id.y, -ends_with("label.y")) %>%
  mutate(totalprop = Count/ TotalCount)

#total number of tracts that doesn't have Asian Alone or in combination   
sum(is.na(mergedata$Count))

submergedata <- mergedata %>%
  filter(POPGROUP.id != "012" & POPGROUP.id != "031") 

#barchart table ready
barpop <- submergedata %>%
  mutate(Asiansubgroup = `POPGROUP.display-label`) %>%
  group_by(POPGROUP.id, Asiansubgroup) %>% #join all the tract population by same subgroup
  summarise(subpoptotal = sum(Count)) %>% 
  mutate(subgroup_prop = round(subpoptotal/sum(data_subset_total$TotalCount), 5))#new column with the proportion of each subgroup population with total population in King County  

#creating barchart 
install.packages("ggplot2") 
library("ggplot2")

ggplot(barpop, aes(x=Asiansubgroup, y=subgroup_prop)) +
  geom_bar(stat="identity", width=.5, fill="tomato3") + 
  labs(title="Proportion of Asian Subgroups with the Total Populatiton in King County", 
        caption="source: Census 2010") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))

# removed first title line from the dataset AsianAlone = AsianAlone[2:nrow(AsianAlone),]

length(unique(AsianAlone$GEO.id2))
#showing the total number of tracts in King County that is populated by Asian groups in the AsianALone dataset

#total population in King County
count(data_subset_total, sum(TotalCount))


```


1. To get a feel for its structure, look at the class, dimensions, column names, structure, and basic summary statistics of your data.




2. Preview the first and last 15 rows of your data. Is you dataset tidy? If not, what principles of tidy data does it seem to be violating?


```{r}
head(data_subset_chinese, 15) 
tail(data_subset_chinese, 15)


```

3. Create a histogram for at least two variables you plan to focus on for your study. Describe what these plots show you about these variables. 

```{r}
AsianAlone <- AsianAlone[-1,]
prop_data <- AsianAlone %>%
  mutate(Count = as.numeric(as.character(D001)), 
         POPGROUP.id = as.numeric(as.character(POPGROUP.id))) %>% # Make new count variable which is numeric version of D001
  
  group_by(POPGROUP.id) %>% # Group by ethnicity/subgroup
  summarise(Total_pop = sum(Count)) %>% #Sum within subgroup
  filter(POPGROUP.id != 031) %>% #Remove asian and other
  mutate(Total_Asian = sum(Total_pop), #Create total asian pop variable
         proportion = Total_pop/Total_Asian) # Creating proportion of subgroup of total Asians
  
hist(data_subset_cambodian$Count)
#Changing factors to numeric on D001 column/population

pop.chinese <- sum(as.numeric(data_subset_chinese$Count), na.rm=FALSE)
pop.cambodian <- sum(as.numeric(data_subset_cambodian$Count), na.rm=FALSE) 
pop.japanese <- sum(as.numeric(data_subset_japanese$Count), na.rm=FALSE) 
pop.korean <- sum(as.numeric(data_subset_korean$Count), na.rm=FALSE)
pop.laotian <- sum(as.numeric(data_subset_laotian$Count), na.rm=FALSE)
pop.hmong <- sum(as.numeric(data_subset_hmong$Count), na.rm=FALSE)
pop.taiwanese <- sum(as.numeric(data_subset_taiwanese$Count), na.rm=FALSE) 
pop.vietnames <- sum(as.numeric(data_subset_vietnamese$Count), na.rm=FALSE)
pop.filipino <- sum(as.numeric(data_subset_filipino$Count), na.rm=FALSE)


pop <- c(pop.chinese, pop.cambodian, pop.japanese, pop.korean, pop.laotian,
                 pop.hmong, pop.taiwanese, pop.vietnames, pop.filipino) %>% as.data.frame() %>% t()

colnames(pop) <- c("Chinese", "Cambodian", "Japanese", "Korean", "Laotian",
                   "Hmong", "Taiwanese", "Vietnames", "Filipino")

```



4. Create at least one bivariate plot showing the relationship between two variables of interest. What does/do the(se) plot(s) tell you about the association between these two variables?

```{r}

plot(prop_data$Total_pop, prop_data$proportion)

# The plot on x-axis is showing the total population by Asian subgroup, and on y-axis is showing the proportion of the group in relationship with the total Asian population. Each circle represents an Asian subgroup.
```


5. Load the `tidyr` package. Do all of your columns correspond to variables? Do any columns represent multiple variables? If your answer is yes to either question, carry out the appropriate `tidyr` function (`gather()` or `spread()` respectively) to tidy your data. 

```{r}
library(tidyr)
# Dataset prop_data is already tidy with columns corresponding group id(ethnic group) count, population count, total Asian population and proportionality of the subgroup within the total Asian population in King County

```




6. Do any columns need to be separated into two or more? Do any columns need to be combined into one? If so, carry out the appropriate the appropriate `tidyr` function (`separate()` or `unite()` respectively) to tidy your data.  


**At this stage each row in your data should represent one observation, each column should be a variable, and each table should be observational unit.** 

7. What is the class `les in your analysis? Are these classes appropriate for the type of measurement they purport to capture? Explain your reasoning.

Variables are appriate as they capture the intended variables of Asian subgroups and proportionality of each subgroup in relationship to the Total asian population. 



8. Do any of your variables need to be coerced into a different data type? If so, carry out the appropriate coercion methods below. (This includes transformation of any date objects using the `lubridate` package)

```{r}
prop_data <- AsianAlone %>%
  mutate(Count = as.numeric(as.character(D001)), 
         POPGROUP.id = as.numeric(as.character(POPGROUP.id))) # Make new count variable which is numeric version of D001
  
#  Data frame variables of proportin and Total population count in prop-data table have been formated from factors to numeric data type, and appropriate to be manipulated as the descriptive analysis will be based on comparing the population subgroups proportionality/disproportionality
  
```



9. Are there any strings you need to manipulate for your analysis? If so, use the appropriate function from the [`stringr`](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html) package. 



10. Do you have any missing values in your dataset? How many and how are they coded? **Be sure to look out for specific codebook values for missing values (i.e. `-1` for `NA`) as well as empty strings or other software-specific values for `NA`.** Don't worry about removing NAs yet - we'll tackle this question later once discern whether they're random or systematically distributed. 

There in no missing values in my data

11. Are there any special values in your dataset? If so, what are they and how do you think they got there? *The presence of special values is less likely if you haven't performed any data manipulation yet so you should remember to return to this step each time you carry out a mathematical transformation of any values in your dataset.*


12. Create a boxplot of your data (you can create an individual boxplot for each variable if there are too many variables in your dataset to meaningfully visualize them all in one plot). Are there any outliers? If so, what are they and to which variable do they correspond? Do any of these outliers seem like obvious errors? If so, why? 


```{r}

Asian_plot <- as.numeric(as.character(AsianAlone$D001))
boxplot(Asian_plot, data=001)

library(ggplot2)
  ggplot(barpop, aes(x=Subgroups, y=Population Proportion)) 
  

#The boxplot shows a great numbers of outliers. The median number or individuals in each tract is less than 500, with many outliers falling outside of both interquartiles, above 1500 head count. An added information, from the 3 

save.image("lab8.RData")

```


13. For any outliers and/or obvious errors, what do you think is the best way to handle them (i.e. remove them entirely, run analyses including and excluding them and compare the results, manually change them to an appropriate measure of center, or something else?). 

In my analysis, it is actually important to consider the outliers as I am trying to represent minority groups within the Asian American community in King County.



