---
title: "Lab 10 - Merging Data"
author: "Your Name Here"
date: "November 2, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r}
setwd("~/Desktop/Undergrad/STAT/Labs")
source(file = "Datasets.R")

library(dplyr)
library(tidyverse)
AsianAlone <- read.csv("DEC_10_SF2_PCT1_with_ann.csv")

Totalpopulation <- read.csv("../Datasets/TotalPopulation/ACS_10_5YR_B01003_with_ann.csv",skip=0)

data_subset_total <- Totalpopulation %>% 
  filter(GEO.id != "Id") %>%
  mutate(TotalCount = as.numeric(as.character(HD01_VD01))) 


#transform from factor to numeric
Data_Asian_Alone <- AsianAlone %>%
  filter(GEO.id != "Id") %>%
  filter(POPGROUP.id != "012" & POPGROUP.id != "031") %>% #removing the Asian Alone and Asian in combination groups
  mutate(Count = as.numeric(as.character(D001))) #creating a new column for the population number as numeric vs character

#barchart
mergedata <- data_subset_total %>% 
  left_join(Data_Asian_Alone, by="GEO.id2") %>%
  select(-GEO.id.y, -ends_with("label.y")) %>%
  mutate(totalprop = Count/ TotalCount)

```

Using your own dataset (which may include more than one table) carry out the following data cleaning steps. Knit together the PDF document and commit both the Lab 10 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository. 

1. For your poster project, do you have multiple tables you'd like to join together to create your complete dataset? If so, describe what each table represents. 


For the poster project, I have created a final dataset for the barchart representing the Asian Subgroups in the x-axis and the propotional number of each subgroup with the total population in King County. To create this table, I have joined the AsianAlone dataset and total population in King County, and divided the numbers creating a new table with the proportion number by each subgroup. For the second visualization, the representation of each subgroup by tract in King County, I have created a table for each subgroup with its tract number/Census, group description and population. At the moment, I am working on how to convey this information visually within my technical skills. The facet maps initial idea might be visually the best way to convey this information, but might not be possible to achieve due to lack of skills. I am currently playing with ggplot2 generating scatterplots but so far haven't found a good visualization. It has been very challenging to convey the representation in each tract, particularly the tracts that does not have any Asian representation. 

2. What is/are your primary key(s)? If you have more than one table in your data, what is/are your foreign key(s)? Do your primary key(s) and foreign key(s) have the same name? If not, what does this mean for the way you need to specify potential data merges?


Primary keys from Census dataset mostly remain. So far, I have renamed and created a foreign key to convey the population proportionality in the barchart visualization. For each group information, it has been created with selecting only existing rows in the same key/subgroup, so no need to mutat and create new columns.



3. If you do not need to merge tables to create your final dataset, create a new dataset from your original dataset with a `grouped_by()` summary of your choice. You will use this separate dataset to complete the following exercises. 

I did merge tables.

If you are merging separate tables as part of your data manipulation process, are your keys of the same data type? If not, what are the differences? Figure out the appropriate coercion process(es) and carry out the steps below. 



4. Perform each version of the mutating joins (don't forget to specify the `by` argument) and print the results to the console. Describe what each join did to your datasets and what the resulting data table looks like. For those joining two separate datasets, did any of these joins result in your desired final dataset? Why or why not?

```{r}
#From lab 8
#barchart table ready
mergedata <- data_subset_total %>% 
  left_join(Data_Asian_Alone, by="GEO.id2") %>%
  select(-GEO.id.y, -ends_with("label.y")) %>%
  mutate(totalprop = Count/ TotalCount)


barpop <- mergedata %>%
  mutate(Asiansubgroup = POPGROUP.display.label) %>%
  group_by(POPGROUP.id, Asiansubgroup) %>%
  summarise(subpoptotal = sum(Count)) %>%
  mutate(subgroup_prop = round(subpoptotal/sum(data_subset_total$TotalCount), 5)) %>%
  filter(POPGROUP.id != "031", POPGROUP.id != "012"
         )
```


5. Do the same thing with the filtering joins. What was the result? Give an example of a case in which a `semi_join()` or an `anti_join()` might be used with your primary dataset

```{r}
alltracts <- inner_join(AsianAlone, Totalpopulation)

```


6. What happens when you apply the set operations joins to your tables? Are these functions useful for you for this project? Explain why or why not. If not, give an example in which one of them might be usefully applied to your data. 

They are useful as the AsianAlone dataset didn't provide information for the tracts that there was no Asians. The missing information is significant for my second visualization as I intend to specifically convey, among other things, the tracts that Asians are not represented in King County.

7. If you have any reason to compare tables, apply `setequal()` below. What were the results? 



8. What is the purpose of binding data and why might you need to take extra precaution when carrying out this specific form of data merging? If your data requires any binding, carry out the steps below and describe what was accomplished by your merge.

```{r}
#barchart table ready
barpop <- mergedata %>%
  mutate(Asiansubgroup = POPGROUP.display.label) %>%
  group_by(POPGROUP.id, Asiansubgroup) %>%
  summarise(subpoptotal = sum(Count)) %>%
  mutate(subgroup_prop = round(subpoptotal/sum(data_subset_total$TotalCount), 5)) %>%
  filter(POPGROUP.id != "031", POPGROUP.id != "012"
  )

#To create the table for the barchart, I had to be careful to specifically select the keys needed to mutate and generate the needed information of the population proportionality, versus merging all the tables. And filter the specific groups (31 and 12) of Asian in combination and Asian Alone subgroups and leaving out of the analysis as it in not clear if these groups are mutually exclusive from the other subgroups population. 



```


9. Do you need to merge multiple tables together using the same type of merge? If so, utilize the `reduce()` function from the `purrr` package to carry out the appropriate merge below. 

I didn't need to use this function to generate the new table.


10. Are there any other steps you need to carry out to further clean, transform, or merge your data into one, final, tidy dataset? If so, describe what they are and carry them out below.


```{r}
# At the moment it has been very challenging to generate a table that would best convey visually the subgroup/tracts representation without using a map. This is the best representation so far, utilizing scatterplots. The challenge with this scatterplot is that with roughly 300 tracts the plots became dwarfed and not as meaningful as I hoped, and the tracts that have no Asian representation from the 9 subgroups are not represented either. I have been spending a lot of time at CSSSR in the past week, a couple of hours at a time, and it has been challenging for them as well. In progress. 


as.numeric(AsianAlone$POPGROUP.display.label)
  
plot(y=as.numeric(AsianAlone$POPGROUP.display.label), x=as.numeric(AsianAlone$GEO.display.label), pch=".")
plot(sort(table(AsianAlone$GEO.display.label[-1])))

```










